---
title: "Geography 176A"
author: "[Chelsey Su](https://suchaoxi.github.io)"
subtitle: "Lab 03: Distances and the Border Zone"
output: 
  html_document:
    theme: journal
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(units)
library(USAboundaries)
library(rnaturalearth)
library(gghighlight)
library(ggrepel)
library(knitr)
library(dplyr)
library(ggthemes)
library(kableExtra)
```

## answering Q1:
```{r}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```
#### info included:
- proj=eqdc -> map projection: equidistant
- lat_0=40 -> latitude of false origin: 
- lon_0=-96 -> longitude: -96
- lat_1=20 -> latitude of first standard parallel: 20
- lat_2=60 -> latitude of second standard parallel: 60
- x_0=0 -> false origin easting: 0
- y_0=0 -> false origin northing: 0
- datum=NAD83 -> datum: NAD83
- units=m -> unit: meter

```{r}
conus = USAboundaries::us_states(resolution = "low") %>% 
  filter(!name %in% c("Alaska", "Hawaii", "Puerto Rico")) %>% 
  st_transform(eqdc)
```

```{r}
countries = countries110 %>% 
  st_as_sf() %>% 
  filter(admin %in% c("Canada", "United States of America", "Mexico")) %>% 
  st_transform(eqdc)
```

```{r, message = FALSE, warning = FALSE}
cities = read_csv("../data/uscities.csv") %>% 
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>% 
  st_transform(eqdc) %>% 
  filter(!state_name %in% c("Alaska", "Hawaii", "Puerto Rico"))
```

## Answering Q2:

### 2.1 - Distance to USA Border (coastline or national) (km)
```{r, warning = FALSE, message = FALSE}
country_border = st_union(conus) %>% 
  st_cast("MULTILINESTRING")

city_boundry = cities %>% 
  mutate(dist_to_border = st_distance(cities, country_border), dist_to_border = set_units(dist_to_border, "km"), dist_to_border = drop_units(dist_to_border))

max5_dist_to_border = city_boundry %>% 
  select(city, state_name, dist_to_border) %>% 
  slice_max(dist_to_border, n = 5) %>% 
  st_drop_geometry()
```

```{r, warning = FALSE, message = FALSE}
kable(max5_dist_to_border, caption = "Five Cities Farthest from the US Border", col.names = c("City", "State", "Distance (km)"), format.args = list(big.mark = ",")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 11)
```
### 2.2 - Distance to States (km)
```{r}
state_border = st_combine(conus) %>% 
  st_cast("MULTILINESTRING")

city_boundry_to_state = cities %>% 
  mutate(dist_to_border = st_distance(cities, state_border), dist_to_border = set_units(dist_to_border, "km"), dist_to_border = drop_units(dist_to_border))

max5_dist_to_state_border = city_boundry_to_state %>% 
  select(city, state_name, dist_to_border) %>% 
  slice_max(dist_to_border, n = 5) %>% 
  st_drop_geometry()
```

```{r}
kable(max5_dist_to_border, caption = "Five Cities Farthest from the State Border", col.names = c("City", "State", "Distance (km)"), format.args = list(big.mark = ",")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 11)
```

### 2.3 - Distance to Mexico (km)
```{r}
mexico_border = countries %>% 
  filter(admin == "Mexico") %>% 
  st_union() %>% 
  st_cast("MULTILINESTRING")

city_boundary_mexico <- cities %>%
    mutate(dist_to_mexico = st_distance(cities, mexico_border), dist_to_mexico = set_units(dist_to_mexico, "km"), dist_to_mexico = drop_units(dist_to_mexico))

max5_to_mexico = city_boundary_mexico %>% 
  select(city, state_name, dist_to_mexico) %>% 
  slice_max(dist_to_mexico, n = 5) %>%
  st_drop_geometry()
```

```{r}
kable(max5_to_mexico, caption = "Five Cities Farthest from Mexican Border", col.names = c("City", "State", "Distance (km)"), format.args = list(big.mark = ",")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 11)
```


### 2.4 - Distance to Canada (km)
```{r}
canada_border = countries %>% 
  filter(admin == "Canada") %>% 
  st_union() %>% 
  st_cast("MULTILINESTRING")

city_boundary_canada <- cities %>%
    mutate(dist_to_canada =st_distance(cities, canada_border), dist_to_canada = set_units(dist_to_canada, "km"), dist_to_canada = drop_units(dist_to_canada))

max5_to_canada = city_boundary_canada %>% 
  select(city, state_name, dist_to_canada) %>% 
  slice_max(dist_to_canada, n = 5) %>%
  st_drop_geometry()
```

```{r}
kable(max5_to_canada, caption = "Five Cities Farthest from Canadian Border", col.names = c("City", "State", "Distance (km)"), format.args = list(big.mark = ",")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 11)
```
## Answering Q3:

### 3.1 Data
```{r}
big_cities = cities %>% 
  slice_max(population, n = 10)

ggplot(big_cities) +
  geom_sf(data = countries) +
  geom_sf(data = conus, color = '#fcba03', lty = 1, size = .5) +
  geom_sf(data = big_cities, size = 2) +
  ggthemes::theme_clean() +
  geom_label_repel(data = big_cities, aes(label = city, geometry = geometry), stat = 'sf_coordinates', size = 2) +
  labs(title = "10 Most Populated Cities in the United States", x = "Longitude", y = "Latitude")
```

### 3.2 City Distance from the Border
```{r}
far_us_border = city_boundry %>% 
  slice_max(dist_to_border, n = 5)

ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = city_boundry, aes(col = dist_to_border), size = 0.25) +
  geom_sf(data = far_us_border) +
  scale_color_gradient(low = "#e6d29c", high = "red") +
  theme_clean() +
  geom_label_repel(data = far_us_border, aes(label = city, geometry = geometry), stat = "sf_coordinates", size = 2) +
  labs(col = "Distance to Border (km)", title = "US Cities and their Distance to US Border", subtitle = "5 Farthest Cities by Distance to US Border Labeled", x = "Longitude", y = "Latitude")
```

### 3.3 City Distance from Nearest State
```{r}
nearest_border = city_boundry_to_state %>% 
  slice_max(dist_to_border, n = 5) 

ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = city_boundry_to_state, aes(col = dist_to_border), size = .25) +
  geom_sf(data = nearest_border) +
  scale_color_gradient(low = "#e6d29c", high = "red") +
  theme_clean() +
  geom_label_repel(data = far_us_border, aes(label = city, geometry = geometry), stat = "sf_coordinates", size = 2) +
  labs(col = "Distance to Border (km)", title = "US Cities and their Distance to State Border", subtitle = "5 Farthest Cities from Nearest State Border Labeled", x = "Longitude", y = "Latitude")
```

### 3.4 Equidistance boundary from Mexico and Canada
```{r}
equi_boundary = cities %>%
  mutate(dist_to_canada = {st_distance(cities, canada_border) %>% set_units('km') %>% drop_units()}, dist_to_mexico = {st_distance(cities, mexico_border) %>% set_units('km') %>% drop_units()}, equidistance = {abs(dist_to_canada - dist_to_mexico)}) %>% 
  filter(equidistance <= 100)

equi_pop = equi_boundary %>% 
  slice_max(population, n = 5)

ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = equi_boundary, size = .5, color = "#e0c655", alpha = 1) +
  geom_sf(data = equi_pop) +
  theme_clean() +
  geom_label_repel(data = equi_pop, aes(label = city, geometry = geometry), stat = "sf_coordinates", size = 4) +
  labs(col = "Distance to Border (km)", title = "US Cities Equidistant to Canada and Mexico Border ± 100 km", subtitle = "Top 5 Most Populated Labeled", x = "Longitude", y = "Latitude")

  
```

## Answering Q4:

### 4.1 Quantifing Border Zone

```{r}
zone_100 = city_boundry %>% 
  filter(dist_to_border <= 160)

number_zone =  nrow(zone_100)

pop_zone_100 = sum(zone_100 $ population)

prop_zone_100 = 100*(pop_zone_100 / sum(city_boundry $ population))

description = c("Number of cities in 100-mile zone", "Population in 100-mile zone", "Proportion of population in 100-mile zone")
value = c("number_zone", "pop_zone_100", "prop_zone_100")

table = data.frame(description, value)


kable(table, caption = '100-Mile Zone Facts', 
      col.names = c("Facts",'Value'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 11)
```


### 4.2 Mapping Border Zone

```{r}
top10_per_state = zone_100 %>% 
  group_by(state_name) %>%
  slice_max(population, n = 1)

ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = city_boundry, aes(col = dist_to_border), size = .1) +
  geom_sf(data = nearest_border) +
  geom_sf(data = top10_per_state) +
  scale_color_gradient(low = "orange", high = "darkred") +
  gghighlight(dist_to_border < 160) +
  theme_clean() +
  geom_label_repel(data = nearest_border,
                   aes(label = city, geometry = geometry),
                   stat = "sf_coordinates",
                   size = 3) +
  labs(col = "Distance to Border (km)", title = "US Cities within 100-miles to US Border", subtitle = "Cities with Distance under 100-miles to US Border Labeled", x = "Longitude", y = "Latitude")
```







